---
title: "Marginal Effects Examples"
author: "Jeffrey Arnold"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Marginal Effects Examples}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r message = FALSE,results='hide'}
library("marfx")
library("dplyr")
library("ggplot2")
library("purrr")
```

## Conflicing Interests and Trade

Example from Kastner (2007) replicated in Berry et. al (2012).

$$
\begin{aligned}[t]
\text{Trade} &= \beta_0 + \beta_C \text{Conflict} + \beta_B \text{(Trade Barriers)}  \\
& + \beta_{BC} (\text{(Trade Barriers)} \times \text{Conflict}) \\
& + \beta Z + \varepsilon 
\end{aligned}
$$



## Xenophobic Voting in Russia

Alexseev (2006) ``Ballot-Box Vigilantism: Ethnic
Population Shifts and Xenophobic Voting in Post-Soviet
Russia.'' *Political Behavior* replicated in Berry et. al (2012).

Alexseev (2006) examines how changes in ethnic composition of Russian regions affects the vote share of the Russian nationalist Zhirinovsky Bloc in the 2003 elections of the Russian State Duma.

The model is,
$$
\begin{aligned}[t]
\text{(Xenophobic Voing)} &= \beta_0 + \beta_{S} \text{(Slavic Share)} \\
&+ \beta_{N} \Delta \text{(non-Slavic Share)} \\
&+ \beta_{SN} (\text{(Slavic Share)} \times \Delta \text{(non-Slavic Share)} \\
&+ \beta Z + \varepsilon
\end{aligned}
$$
where $Z$ are control variables.
```{r}
mod1 <- lm(xenovote ~ slavicshare * changenonslav +
     inc9903 + eduhi02 + unemp02 + apt9200 + vsall03 + brdcont, 
   data = alexseev)
```
Note that even though the `alexseev` dataset includes a variable
`slavicshare_changenonslav` which is the interaction between Slavic share and $\Delta$ non-Slavic share, this model specifies the interaction in the formula. This makes it much simpler for analysis since the `lm` object will contain information about the relationship between the variables and the terms in the regression, so in the analysis we will only need to specify changes in `slavicshare` or `changenonslav` and the functions can infer that the interaction will need to change.  If instead, the varible `slavicshare_changenonslav` was used, the `lm` object would not know that the variable was an interaction between `slavicshare` and `changenonslav` and we would need to be careful to always change both the variable of interest and the interaction when considering any marginal or partial effects.

Berry et. al (2012, p. 14) present two marginal effect plots.
The marginal effect of Slavic share on xenophobic voting, and the marginal effect of $\Delta$ non-Slavic share on xenophobic voting.
```{r fig.caption="Marginal Effect of Slavic Share on Xenophobic Voting"}
.data <- cross_d(list(changenonslav =
                   seq(min(alexseev$changenonslav),
                       max(alexseev$changenonslav),
                       length.out = 30),
                 slavicshare = mean(alexseev$slavicshare),
                 inc9903 = mean(alexseev$inc9903),
                 eduhi02 = mean(alexseev$eduhi02),
                 unemp02 = mean(alexseev$unemp02),
                 apt9200 = mean(alexseev$apt9200),
                 vsall03 = mean(alexseev$vsall03),
                 brdcont = FALSE))

.data <- cbind(.data, mfx(mod1, data = .data, "slavicshare"))

ggplot() +
  geom_ribbon(data = .data, mapping = aes(x = changenonslav, 
                                          ymin = conf.low, 
                                          ymax = conf.high),
              alpha = 0.3) +
  geom_line(data = .data, mapping = aes(x = changenonslav, y = estimate)) +
  geom_rug(data = alexseev, mapping = aes(x = changenonslav)) +
  geom_hline(yintercept = 0, color = "gray") + 
  scale_x_continuous(name = expression(paste(Delta, "non-Slavic Share")), breaks = seq(-2, 12, by = 2)) +
  scale_y_continuous(name = "Marginal Effect of Slavic Share",
                     breaks = seq(-.05, .25, by = .05)) +
  theme_minimal() +
  theme(panel.grid = element_blank())
     
```
The marginal effects of $\Delta$ non-Slavic Share on Xenophobic voting:
```{r}
.data <- cross_d(list(slavicshare =
                   seq(min(alexseev$slavicshare),
                       max(alexseev$slavicshare),
                       length.out = 30),
                 changenonslav = mean(alexseev$changenonslav),
                 inc9903 = mean(alexseev$inc9903),
                 eduhi02 = mean(alexseev$eduhi02),
                 unemp02 = mean(alexseev$unemp02),
                 apt9200 = mean(alexseev$apt9200),
                 vsall03 = mean(alexseev$vsall03),
                 brdcont = FALSE))

.data <- cbind(.data, mfx(mod1, data = .data, "changenonslav"))

ggplot() +
  geom_ribbon(data = .data, mapping = aes(x = slavicshare,
                                          ymin = estimate - 1.96 * std.error, 
                                          ymax = estimate + 1.96 * std.error),
              alpha = 0.3) +
  geom_line(data = .data, mapping = aes(x = slavicshare, y = estimate)) +
  geom_rug(data = alexseev, mapping = aes(x = slavicshare)) +
  geom_hline(yintercept = 0, color = "gray") + 
  scale_x_continuous(name = "Slavic Share", breaks = seq(30, 100, by = 10)) +
  scale_y_continuous(name = expression(paste("Marginal Effect of", Delta, "non-Slavic Share")),
                     breaks = seq(-1, .5, by = .25)) +
  theme_minimal() +
  theme(panel.grid = element_blank())
```
By using the `std.error`, this plot has smoother confidence intervals than the one using `conf.low` and `conf.high`, which are generated from the quantiles of the simulations.


**TODO:** Add clustered standard errors.

To summarize the effect of Slavic Share and $\Delta$ non-Slavic Share on xenophobic vote share we can calculate the average marginal effects of these variables,
```{r fig.caption="Average Marginal Effects on Xenophobic Voting"}
variables <- c("slavicshare" = "Slavic Share",
               "changenonslav" = "Change non-Slavic Share")
.data <- list()
for (i in seq_along(variables)) {
  .data[[i]] <-
    amfx(mod1, names(variables)[i], data = alexseev) %>%
    mutate(variable = names(variables)[i],
           description = variables[i])
}
.data <- bind_rows(.data)

ggplot(.data, aes(x = description, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_hline(yintercept = 0, colour = "gray") +
  geom_pointrange() +
  coord_flip() +
  scale_x_discrete("") +
  scale_y_continuous("Average Marginal Effects") +
  theme_minimal() +
  theme(panel.grid = element_blank())
```

**TODO** What to do if interactions are specified in the 
data rather than the formula.

## Logit

A logit model of voting turnout.

```{r}
library("Zelig")
data("turnout", package = "Zelig")
mod_turnout <- glm(vote ~ poly(age, 2) + race + educate, data = turnout, family = binomial)
summary(mod_turnout)
```

Calculate the average marginal effect of age,
```{r}
amfx(mod_turnout, "age", data = turnout)
```
the average marginal effect of all variables,
```{r}
data.frame(variable = c("age", "race", "educate")) %>%
  group_by(variable) %>%
  do({
    amfx(mod_turnout, .$variable, data = turnout)
  })

```

Calculate the average marginal effects of age at its mean value, holding the other
values at their means and race = "white",
```{r}
summarize(turnout, age = mean(age), educate = mean(educate)) %>%
  mutate(race = factor("white", levels = levels(turnout$race))) %>%
  mfx(mod_turnout, "age", data = .)
```
Calculate the marginal effects of age at each decade, 20 to 80, holding `educate` at its mean,
and setting `race` to "white":
```{r}
data_frame(age = seq(20, 80, by = 10)) %>%
  mutate(race = factor("white", levels = levels(turnout$race)),
         educate = mean(turnout$educate)) %>%
  bind_cols(., mfx(mod_turnout, "age", data = .)) %>%
  ggplot(aes(x = age, y = estimate,
             ymin = estimate - 2 * std.error, ymax = estimate + 2 * std.error)) +
  geom_ribbon(alpha = 0.3) +
  geom_line()
```


Calculate the average marginal effects of each variable holding the others at
representative values within each group
```{r}
mfx(mod_turnout, "age", data = summarize(group_by(turnout, race),
                                         age = mean(age),
                                         educate = mean(educate)))
```



